---
title: "WorkingCode"
author: "Douglas Hannum"
date: "4/11/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(lme4)
library(dplyr)
library(MuMIn)
```

#Data Setup

Loading the data and getting rid of subject 16 because they had tons of missing data.  Also I am not using any observations that do not have a recorded ingress strategy.

```{r loading data}
# Loading data
df <- read.csv('./data/merged_data.csv', header = T)
# Getting rid of subject 16
df <- df[df$Subject != 'S16',]
# Making sure all observations have an ingress strategy
df <- df[df$Ingress.Strategy != '',]
dim(df)
length(unique(df$Subject))
```

We now have 637 observations for 47 subjects.

#Basic Plots of Ingress

Looking at ingress time for different vehicles with subsets for group.  Another one looking at the ingress time by vehicle adjusting for vehicle order.
```{r plot ingress}
ggplot (df, aes(x = Vehicle.and.Door.Condition, y = IngressTime..s.,
                     fill = Group)) +
        geom_boxplot()

df$Vehicle.Order <- as.factor(df$Vehicle.Order)
ggplot (df, aes(x = Vehicle.and.Door.Condition, y = IngressTime..s.,
                     fill = Vehicle.Order)) +
        geom_boxplot()
```

#Summarizing Physical Measurements

Creating three different summary statistics for one-leg balance.  Standardizeing both age and weight.  Separating the vehicle and door condition variable into two different variables.  

```{r physical summaries}
#colnames(df)
df$first <- T
for (i in 2:dim(df)[1]){
        df$first[i] <- df$Subject[i] != df$Subject[i-1]
}
#summary(df$first)
olb <- colnames(df)[49:54]
#olb
#summary(df[,c(olb)])
#class(df[,49])
df$olb_max <- NULL
df$olb_mean_r <- NULL
df$olb_mean_l <- NULL
for (i in 1:dim(df)[1]){
        df$olb_max[i] <- max(df[i,c(olb)])
        df$olb_mean_l[i] <- mean(c(df$OLB_L1[i], df$OLB_L2[i], df$OLB_L3[i]),
                                 na.rm = T)
        df$olb_mean_r[i] <- mean(c(df$OLB_R1[i], df$OLB_R2[i], df$OLB_R3[i]),
                                 na.rm = T)
}
#class(df$olb_max)
olbs <- df[df$first == T,]
df$olb_mean_sum <- df$olb_mean_l + df$olb_mean_r

ggplot(data = df[df$first == T,], aes(x = olb_mean_sum)) +
        geom_histogram(binwidth = 2) + 
        geom_vline(xintercept = 15)
df$OLB_Rating <- NA
for (i in 1:dim(df)[1]){
        if (is.na(df$olb_mean_sum[i])){
                df$OLB_Rating[i] <- NA
        }
        else if (df$olb_mean_sum[i] == 60){
                df$OLB_Rating[i] <- 3
        }
        else if (df$olb_mean_sum[i] > 15){
                df$OLB_Rating[i] <- 2
        }
        else {
                df$OLB_Rating[i] <- 1
        }
}
df$OLB_Rating <- as.factor(df$OLB_Rating)

df$Age <- df$Age..yrs..x
#summary(df[df$first == T,]$Age)
df$sd_age <- NA
for (i in 1:dim(df)[1]){
        df$sd_age[i] <- (df$Age[i] - 76.64) / 7
}

df <- separate(df, Vehicle.and.Door.Condition, 
               into = c('Vehicle','Door Condition'), sep = '_')

#summary(df[df$first == T,]$Weight.kg.)
#sd(df[df$first == T,]$Weight.kg.)
for (i in 1:dim(df)[1]){
        df$sd_weight[i] <- (df$Weight.kg.[i] - 81.89) / 17.5
}
#summary(df[df$first == T,]$sd_weight)

df$Gender <- df$Gender.x
```

# Looking at the Ingress Strategy

Want to see if people are switching strategy throughout the study.

```{r ingress strategy}
subj <- unique(as.character(df$Subject))
ingress_df <- matrix(nrow = length(subj), ncol = 2)
rownames(ingress_df) <- subj
colnames(ingress_df) <- c('Normal','Two_Feet')
w <- 1
for (i in subj){
        subj_df <- df[df$Subject == i,]
        p <- summary(subj_df$Ingress.Strategy)
        ingress_df[w,] <- p[c(2,3)]
        w <- w + 1
}
#ingress_df
```
Six people were about half and half in strategy (5x 8-6, 1x 6-7) and one person who was a 3-11.  For the most part subjects stayed consistent with their ingress strategy.

```{r ingress time distribution}
ggplot(data =df, aes(x = IngressTime..s.)) + geom_histogram() +
        ggtitle('Distribution of Ingress Time') +
        ylab ('Counts') + xlab('Ingress Time (s)') + theme_bw() +
        theme(plot.title = element_text(hjust = 0.5))
# ggsave('./images/ingress_plot.png', device = 'png', width = 7, height = 4,
#        units = 'in')
```

#Boxplot Images

Creating some boxplot images to visualize differences

```{r boxplot setup}
df2 <- df
df2$Group <- as.character(df2$Group)
for (i in 1:dim(df2)[1]){
        if ((df2$Group[i]) == 'Neuropathy'){
                df2$Group[i] <- ('PN')
        }
}
df2$Group <- as.factor(df2$Group)
#summary(df2$Group)

for (i in 1:dim(df2)[1]){
        if (!is.na(df2$`Door Condition`[i])){
                if(df2$`Door Condition`[i] == 'Pklt'){
                        df2$`Door Condition`[i] <- 'Partially Open'
                }
        }
}


colnames(df2)[36] <- 'Ingress Strategy'
df2$`Ingress Strategy` <- as.character(df2$`Ingress Strategy`)
for (i in 1:dim(df2)[1]){
        if (df2$`Ingress Strategy`[i] == 'twofeet'){
                df2$`Ingress Strategy`[i] <- 'Two Feet'
        }
        else if (df2$`Ingress Strategy`[i] == 'normal'){
                df2$`Ingress Strategy`[i] <- 'Normal'
        }
}
```

```{r boxplots}
ggplot (df2, aes(x = Vehicle, y = IngressTime..s.,, fill = `Door Condition`)) +
        geom_boxplot () + theme_bw() +
        ylab('Ingress Time (s)') + xlab ('Vehicle') + 
        ggtitle('Ingress Time for Each Vehicle', subtitle = 'Split by Door Condition') +
        theme(plot.title = element_text(hjust = 0.5), 
              plot.subtitle = element_text(hjust = 0.5)) +
        scale_fill_discrete(name = "Door Condition")

# ggsave('./images/IngressByVehicleByDoorStatus.png', device = 'png', 
#        height = 4, width = 7, units = 'in')

ggplot (df2, aes(x = Vehicle, y = IngressTime..s.,
                     fill = Group)) +
        geom_boxplot() + theme_bw() +
        ylab ('Ingress Time (s)') + xlab ('Vehicle') +
        ggtitle ('Ingress Time for Each Vehicle', subtitle = 'Split by Group') +
        theme(plot.title = element_text(hjust = 0.5), 
              plot.subtitle = element_text(hjust = 0.5))

# ggsave('./images/IngressByVehicleByGroup.png',device = 'png', height = 4,
#        width = 7, units = 'in')

ggplot (df2[df2$`Ingress Strategy` != '',], aes(x = Vehicle, y = IngressTime..s., 
                                               fill = `Ingress Strategy`)) +
        geom_boxplot() + theme_bw() + ylab ('Ingress Time (s)') +
        xlab ('Vehicle') + 
        ggtitle ('Ingress Time for Each Vehicle', subtitle = 'Split by Strategy') +
        theme(plot.title = element_text(hjust = 0.5), 
              plot.subtitle = element_text(hjust = 0.5)) +
        scale_fill_discrete(name = "Ingress Strategy")

# ggsave('./images/IngressByVehicleByStrategy.png', device = 'png', height = 4,
#        width = 7, units = 'in')
```

#Creating Table for Univariate ME Models

```{r}
df3 <- df2
mo1 <- lmer(data = df3, IngressTime..s. ~ (1|Subject) + `Ingress Strategy`)
mo2 <- lmer(data = df3, IngressTime..s. ~ (1|Subject) + Group)
mo3 <- lmer(data = df3, IngressTime..s. ~ (1|Subject) + OLB_Rating)

tbl1 <- matrix(ncol = 3)
tbl1 <- rbind(tbl1,summary(mo1)$coefficients)
tbl1 <- rbind(tbl1,summary(mo2)$coefficients)
tbl1 <- rbind(tbl1,summary(mo3)$coefficients)
tbl1 <- tbl1[-1,]
tbl1 <- as.data.frame(tbl1)
tbl1$lb <- tbl1$Estimate - 1.96* tbl1$`Std. Error`
tbl1$ub <- tbl1$Estimate + 1.96* tbl1$`Std. Error`
tbl1$CI <- paste0('(', round(tbl1$lb,2), ', ', round(tbl1$ub,2), ')')

# write.csv(tbl1, './tables/Ingress_naive_model.csv', quote = F, row.names = F)
```

We see that ingress strategy is a significant covariate when determining ingress time.

#An Adjusted Model

Adjusting for multiple variables and using random effects for both subject and vehicle.  Then getting confidence intervals and outputting a table.
```{r adj model}
mod1 <- lmer(data = df2, IngressTime..s. ~ (1|Subject) + (1|Vehicle) +
        `Ingress Strategy` + `Door Condition` + Gender + sd_weight +
        sd_age + OLB_Rating + Group)

tbl2 <- as.data.frame(summary(mod1)$coefficients)
tbl2$CI <- paste0('(', round(tbl2$Estimate - 1.96*tbl2$`Std. Error`,2), ', ',
                  round(tbl2$Estimate + 1.96*tbl2$`Std. Error`, 2), ')')
tbl2$lb <- tbl2$Estimate - 1.96* tbl2$`Std. Error`
tbl2$ub <- tbl2$Estimate + 1.96* tbl2$`Std. Error`
tbl2$CI <- paste0('(', round(tbl2$lb,2), ', ', round(tbl2$ub,2), ')')
tbl2
# write.csv(tbl2,'./tables/ingress_adj_model.csv', quote = F, row.names = F)
```
Ingress strategy, door condition, standardized weight and standardized age are significant.


Looking at the model residuals

```{r mod residuals}
res <- summary(mod1)$residuals
plot(res)
rdf <- as.data.frame(res)
ggplot(data = rdf, aes (x = rdf[,1])) + geom_histogram(bins = 7)
```

I think they look pretty good, but I did not test further.


#Aim2

Getting minimums, means and maximums for all the physical variables.

```{r summarizing all variables}
df3 <- df2

df3 <- df3[,c(1,34,43,44,56:83)]
df3$Grip_r_mean <- rowMeans(df3[,5:7], na.rm = T)
df3$GriplefthanMean <- rowMeans(df3[,8:10], na.rm = T)
df3$GripMean <- rowMeans(df3[,5:10], na.rm = T)
df3$KneeExtensionMean <- rowMeans(df3[,13:15], na.rm = T)
df3$HipAbductionMean <- rowMeans(df3[,16:18], na.rm = T)
df3$HipRotMean <- rowMeans(df3[,19:21], na.rm = T)

df3$GriprighthandMax <- apply(df3[,5:7], 1, max)
df3$GriplefthanMax <- apply(df3[,8:10], 1, max)
df3$GripMax <- apply(df3[,5:10], 1, max)
df3$KneeExtensionMax <- apply(df3[,13:15], 1, max)
df3$HipAbductionMax <- apply(df3[,16:18], 1, max)
df3$HipRotationMax <- apply(df3[,19:21], 1, max)

df3$GriprighthandMin <- apply(df3[,5:7], 1, min)
df3$GriplefthanMin <- apply(df3[,8:10], 1, min)
df3$GripMin <- apply(df3[,5:10], 1, min)
df3$KneeExtensionMin <- apply(df3[,13:15], 1, min)
df3$HipAbductionMin <- apply(df3[,16:18], 1, min)
df3$HipRotationMin <- apply(df3[,19:21], 1, min)
```

Creating an Aim2 plot

Running univariate models adjusting for subject (random effect), then getting an r-squared value for each variable.  Then plotting it.  Tested all the variables.  Only used one statistic per variable, which was the one which described the most variation.

```{r aim2 plot}
variables <- names(df3)[-c(1,2, 5:10, 13:21)]
table <- matrix(ncol = 2, nrow = length(variables))
rownames(table) <- variables
colnames(table) <- c('R2m','R2c')
for (i in (variables)){
        form <- paste0('(IngressTime..s.) ~ (1|Subject) + `', i, '`')
        mod <- lmer(data = df3, form)
        table[i,c(1,2)] <- r.squaredGLMM(mod)[c(1,2)]
}
tbl <- as.data.frame(table)

tbl$Variables <- variables
tbl <- tbl[order(tbl$R2m,decreasing = T),]
tbl$`Variance Explained` <- round(tbl$R2m *100,2)
tbl_plot <- tbl[c(1,3,4,6,8,9,10,11,12,13,14),]
tbl_plot$Variables <- c('OLB Rating', 'Group','MDNS','Off Accuracy', 'Age',
                        'Accuracy','RRT', 'Knee Extension Max',
                        'Hip Abduction Mean','MMSE', 'SRT')
ggplot(data = tbl_plot, aes(x = reorder(Variables, `Variance Explained`), 
                            y = `Variance Explained`)) +
        geom_bar(stat = 'identity') + coord_flip() + xlab('Variable') + 
        ylab('% of Variance Explained') +
        ggtitle('Ingress Time Variance Explained', 
                subtitle =  'by Variables, adjusted for Subject') +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5), 
              plot.subtitle = element_text(hjust = 0.5))

# ggsave('./images/Ingress_Aim2Plot.png', device = 'png', height = 4,
#        width = 7, units = 'in')
```

